GPU Efficient Texture Based Bezier Curve Evaluation

### 2015-07-19 Paper/paper.tex
v 0.99 -> basically ready to review and me re-read a few times!

\begin{lstlisting}[caption={GLSL for evaluating a cubic curve encoded in a 4 pixel 1d texture.  Linear texture sampling is used to evaluate the linear level of the De Casteljeau algorithm, then the process is continued both with the De Casteljeau algorithm, as well as the Bernstein form of a quadratic Bezier curve.}, label={lst:GLSLCubicTexture1D}]
uniform sampler1D uSampler;  // 4x1 texture: $P_0,P_1,P_2,P_3$
const float c_textureSize = 4.0;

@@

\begin{lstlisting}[caption={GLSL for evaluating a cubic curve encoded in a $(2,4)$ pixel 2d texture.  Bilinear texture sampling used to evaluate the first two levels of the De Casteljeau algorithm, then the process is continued both with the De Casteljeau algorithm, as well as the Bernstein form of a linear Bezier curve (lerp).}, label={lst:GLSLCubicTexture2D}]
// 2x4 2d texture, rounded to the nearest powers of 2 from 2x3.
uniform sampler2D uSampler; 
const vec2 c_textureSize = vec2(2.0,4.0);

@@

TODO: continue from this point cleaning up notation (overline BS as well as A,B,C control points)

@@

You can combine two cubic curves to make a quartic curve, or three cubic curves to make a quintic curve.  If you were to put $2$x$2$x$2$ cubic curves back to back, you would notice a redundancy like we saw in lower dimensionality textures.  Instead of going that route though, we can just build on what we did for two dimensional textures.

@@

\begin{figure}
    \begin{tikzpicture}
      \draw[step=1cm,gray,very thin] (0,0,0) grid (2,-2,0);
      \draw (0.5,-0.5) node {$A$};
      \draw (1.5,-0.5) node {$B$};
      \draw (0.5,-1.5) node {$B$};
      \draw (1.5,-1.5) node {$C$};

      \begin{scope}[shift={(0.75,0.75,0)}]
        \draw[step=1cm,gray,dashed] (0,0,0) grid (2,-2,0);
        \draw[gray] (0.5,-0.5) node {$B$};
        \draw[gray] (1.5,-0.5) node {$C$};
        \draw[gray] (0.5,-1.5) node {$C$};
        \draw[gray] (1.5,-1.5) node {$D$};
      \end{scope}

      \draw[dashed] (0,0,0) -- (0.75, 0.75,0);
      \draw[dashed] (2,0,0) -- (2.75, 0.75,0);
      \draw[dashed] (0,-2,0) -- (0.75,-1.25,0);
      \draw[dashed] (2,-2,0) -- (2.75,-1.25,0);
    \end{tikzpicture}
    \hspace{5mm}
    \begin{tikzpicture}
      \draw[step=1cm,gray,very thin] (0,0,0) grid (4,-2,0);
      \draw (0.5,-0.5) node {$A$};
      \draw (1.5,-0.5) node {$B$};
      \draw (0.5,-1.5) node {$B$};
      \draw (1.5,-1.5) node {$C$};

      \draw (2.5,-0.5) node {$E$};
      \draw (3.5,-0.5) node {$F$};
      \draw (2.5,-1.5) node {$F$};
      \draw (3.5,-1.5) node {$G$};

      \begin{scope}[shift={(0.75,0.75,0)}]
        \draw[step=1cm,gray,dashed] (0,0,0) grid (4,-2,0);
        \draw[gray] (0.5,-0.5) node {$B$};
        \draw[gray] (1.5,-0.5) node {$C$};
        \draw[gray] (0.5,-1.5) node {$C$};
        \draw[gray] (1.5,-1.5) node {$D$};

        \draw[gray] (2.5,-0.5) node {$F$};
        \draw[gray] (3.5,-0.5) node {$G$};
        \draw[gray] (2.5,-1.5) node {$G$};
        \draw[gray] (3.5,-1.5) node {$H$};
      \end{scope}

@@

    \begin{tikzpicture}
      \draw[step=1cm,gray,very thin] (0,0,0) grid (2,-3,0);
      \draw (0.5,-0.5) node {$A$};
      \draw (1.5,-0.5) node {$B$};
      \draw (0.5,-1.5) node {$B$};
      \draw (1.5,-1.5) node {$C$};
      \draw (0.5,-2.5) node {$C$};
      \draw (1.5,-2.5) node {$D$};

      \begin{scope}[shift={(0.75,0.75,0)}]
        \draw[step=1cm,gray,dashed] (0,0,0) grid (2,-3,0);
        \draw[gray] (0.5,-0.5) node {$B$};
        \draw[gray] (1.5,-0.5) node {$C$};
        \draw[gray] (0.5,-1.5) node {$C$};
        \draw[gray] (1.5,-1.5) node {$D$};
        \draw[gray] (0.5,-2.5) node {$D$};
        \draw[gray] (1.5,-2.5) node {$E$};
      \end{scope}

      \draw[dashed] (0,0,0) -- (0.75, 0.75,0);
      \draw[dashed] (2,0,0) -- (2.75, 0.75,0);
      \draw[dashed] (0,-3,0) -- (0.75,-2.25,0);
      \draw[dashed] (2,-3,0) -- (2.75,-2.25,0);
    \end{tikzpicture}
    \hspace{5mm}
    \begin{tikzpicture}
      \draw[step=1cm,gray,very thin] (0,0,0) grid (4,-3,0);
      \draw (0.5,-0.5) node {$A$};
      \draw (1.5,-0.5) node {$B$};
      \draw (0.5,-1.5) node {$B$};
      \draw (1.5,-1.5) node {$C$};
      \draw (0.5,-2.5) node {$C$};
      \draw (1.5,-2.5) node {$D$};

      \draw (2.5,-0.5) node {$F$};
      \draw (3.5,-0.5) node {$G$};
      \draw (2.5,-1.5) node {$G$};
      \draw (3.5,-1.5) node {$H$};
      \draw (2.5,-2.5) node {$H$};
      \draw (3.5,-2.5) node {$I$};

      \begin{scope}[shift={(0.75,0.75,0)}]
        \draw[step=1cm,gray,dashed] (0,0,0) grid (4,-3,0);
        \draw[gray] (0.5,-0.5) node {$B$};
        \draw[gray] (1.5,-0.5) node {$C$};
        \draw[gray] (0.5,-1.5) node {$C$};
        \draw[gray] (1.5,-1.5) node {$D$};
        \draw[gray] (0.5,-2.5) node {$D$};
        \draw[gray] (1.5,-2.5) node {$E$};

        \draw[gray] (2.5,-0.5) node {$G$};
        \draw[gray] (3.5,-0.5) node {$H$};
        \draw[gray] (2.5,-1.5) node {$H$};
        \draw[gray] (3.5,-1.5) node {$I$};
        \draw[gray] (2.5,-2.5) node {$I$};
        \draw[gray] (3.5,-2.5) node {$J$};
      \end{scope}

@@

For instance, if you had a 2x2 texture, you could encode the control points A,B,C into the red channel, and the control points B,C,D into the green channel.  Then, after you did a texture read, you could do a lerp between the red and the green channel.  Since the red and green channel each contain a quadratic curve, the result would be a cubic curve, defined by the control points A,B,C,D.  However, you've used up two of the channels for a single curve, so are getting fewer curves per texture read, but they are higher degree.  The blue and the alpha channel could either be quadratic curves, or could also be encoded such that they could be combined together to provide another cubic curve.

@@

If you you are combining M color channels, each of which are values from curves of degree N, the resulting curve value will be degree N+M-1.  The intuitive explanation to this, is that each channel has the ability to add one more control point to the curve, regardless of what texture dimensionality you are working with.

An interesting side effect to combining color channels is that the accuracy of the resulting curve gets better.  This is because you are doing some of the math in full precision floating point, instead of the limited precision of the fixed point math in the texture sampler.

@@

%-------------------------------------------------------------------------
\section*{Index of Supplemental Materials}
\label{sec:indexofsupplementalmaterials}
TODO: this

>>>>

\begin{lstlisting}[caption={GLSL for evaluating a cubic curve encoded in a 4 pixel 1d texture.  Linear texture sampling is used to evaluate the linear level of the De Casteljeau algorithm, then the process is continued both with the De Casteljeau algorithm, as well as the Bernstein form of a quadratic Bezier curve.}, label={lst:GLSLCubicTexture1D}]
uniform sampler1D uSampler;  
const float c_textureSize = 4.0; // $P_0,P_1,P_2,P_3$

@@

\begin{lstlisting}[caption={GLSL for evaluating a cubic curve encoded in a $(2,4)$ pixel 2d texture.  Bilinear texture sampling used to evaluate the first two levels of the De Casteljeau algorithm, then the process is continued both with the De Casteljeau algorithm, as well as the Bernstein form of a linear Bezier curve (lerp).}, label={lst:GLSLCubicTexture2D}]
uniform sampler2D uSampler; 
const vec2 c_textureSize = vec2(2.0,4.0); // rounded up from 2x3

@@

@@

You can combine two cubic curves to make a quartic curve, or three cubic curves to make a quintic curve.  If you were to put $2$x$2$x$2$ cubic curves back to back, you would notice a redundancy like we saw in lower dimensionality textures.  Instead of going that route, we can build on what we did for two dimensional textures.

@@

\begin{figure}
    \begin{tikzpicture}
      \draw[step=1cm,gray,very thin] (0,0,0) grid (2,-2,0);
      \draw (0.5,-0.5) node {$P_0$};
      \draw (1.5,-0.5) node {$P_1$};
      \draw (0.5,-1.5) node {$P_1$};
      \draw (1.5,-1.5) node {$P_2$};

      \begin{scope}[shift={(0.75,0.75,0)}]
        \draw[step=1cm,gray,dashed] (0,0,0) grid (2,-2,0);
        \draw[gray] (0.5,-0.5) node {$P_1$};
        \draw[gray] (1.5,-0.5) node {$P_2$};
        \draw[gray] (0.5,-1.5) node {$P_2$};
        \draw[gray] (1.5,-1.5) node {$P_3$};
      \end{scope}

      \draw[dashed] (0,0,0) -- (0.75, 0.75,0);
      \draw[dashed] (2,0,0) -- (2.75, 0.75,0);
      \draw[dashed] (0,-2,0) -- (0.75,-1.25,0);
      \draw[dashed] (2,-2,0) -- (2.75,-1.25,0);
    \end{tikzpicture}
    \hspace{5mm}
    \begin{tikzpicture}
      \draw[step=1cm,gray,very thin] (0,0,0) grid (4,-2,0);
      \draw (0.5,-0.5) node {$P_0$};
      \draw (1.5,-0.5) node {$P_1$};
      \draw (0.5,-1.5) node {$P_1$};
      \draw (1.5,-1.5) node {$P_2$};

      \draw (2.5,-0.5) node {$P_4$};
      \draw (3.5,-0.5) node {$P_5$};
      \draw (2.5,-1.5) node {$P_5$};
      \draw (3.5,-1.5) node {$P_6$};

      \begin{scope}[shift={(0.75,0.75,0)}]
        \draw[step=1cm,gray,dashed] (0,0,0) grid (4,-2,0);
        \draw[gray] (0.5,-0.5) node {$P_1$};
        \draw[gray] (1.5,-0.5) node {$P_2$};
        \draw[gray] (0.5,-1.5) node {$P_2$};
        \draw[gray] (1.5,-1.5) node {$P_3$};

        \draw[gray] (2.5,-0.5) node {$P_5$};
        \draw[gray] (3.5,-0.5) node {$P_6$};
        \draw[gray] (2.5,-1.5) node {$P_6$};
        \draw[gray] (3.5,-1.5) node {$P_7$};
      \end{scope}

@@

    \begin{tikzpicture}
      \draw[step=1cm,gray,very thin] (0,0,0) grid (2,-3,0);
      \draw (0.5,-0.5) node {$P_0$};
      \draw (1.5,-0.5) node {$P_1$};
      \draw (0.5,-1.5) node {$P_1$};
      \draw (1.5,-1.5) node {$P_2$};
      \draw (0.5,-2.5) node {$P_2$};
      \draw (1.5,-2.5) node {$P_3$};

      \begin{scope}[shift={(0.75,0.75,0)}]
        \draw[step=1cm,gray,dashed] (0,0,0) grid (2,-3,0);
        \draw[gray] (0.5,-0.5) node {$P_1$};
        \draw[gray] (1.5,-0.5) node {$P_2$};
        \draw[gray] (0.5,-1.5) node {$P_2$};
        \draw[gray] (1.5,-1.5) node {$P_3$};
        \draw[gray] (0.5,-2.5) node {$P_3$};
        \draw[gray] (1.5,-2.5) node {$P_4$};
      \end{scope}

      \draw[dashed] (0,0,0) -- (0.75, 0.75,0);
      \draw[dashed] (2,0,0) -- (2.75, 0.75,0);
      \draw[dashed] (0,-3,0) -- (0.75,-2.25,0);
      \draw[dashed] (2,-3,0) -- (2.75,-2.25,0);
    \end{tikzpicture}
    \hspace{5mm}
    \begin{tikzpicture}
      \draw[step=1cm,gray,very thin] (0,0,0) grid (4,-3,0);
      \draw (0.5,-0.5) node {$P_0$};
      \draw (1.5,-0.5) node {$P_1$};
      \draw (0.5,-1.5) node {$P_1$};
      \draw (1.5,-1.5) node {$P_2$};
      \draw (0.5,-2.5) node {$P_2$};
      \draw (1.5,-2.5) node {$P_3$};

      \draw (2.5,-0.5) node {$P_5$};
      \draw (3.5,-0.5) node {$P_6$};
      \draw (2.5,-1.5) node {$P_6$};
      \draw (3.5,-1.5) node {$P_7$};
      \draw (2.5,-2.5) node {$P_7$};
      \draw (3.5,-2.5) node {$P_8$};

      \begin{scope}[shift={(0.75,0.75,0)}]
        \draw[step=1cm,gray,dashed] (0,0,0) grid (4,-3,0);
        \draw[gray] (0.5,-0.5) node {$P_1$};
        \draw[gray] (1.5,-0.5) node {$P_2$};
        \draw[gray] (0.5,-1.5) node {$P_2$};
        \draw[gray] (1.5,-1.5) node {$P_3$};
        \draw[gray] (0.5,-2.5) node {$P_3$};
        \draw[gray] (1.5,-2.5) node {$P_4$};

        \draw[gray] (2.5,-0.5) node {$P_6$};
        \draw[gray] (3.5,-0.5) node {$P_7$};
        \draw[gray] (2.5,-1.5) node {$P_7$};
        \draw[gray] (3.5,-1.5) node {$P_8$};
        \draw[gray] (2.5,-2.5) node {$P_8$};
        \draw[gray] (3.5,-2.5) node {$P_9$};
      \end{scope}

@@

For instance, if you had a 2x2 texture, you could encode the control points $P_0,P_1,P_2$ into the red channel, and the control points $P_1,P_2,P_3$ into the green channel.  Then, after you did a texture read, you could do a lerp between the red and the green channel values.  Since the red and green channel each contain a quadratic curve, the result would be a cubic curve, defined by the control points $P_0,P_1,P_2,P_3$  However, you've used up two of the channels for a single curve, so are getting fewer curves per texture read, but they are higher degree.  The blue and the alpha channel could either be quadratic curves, or could also be encoded such that they could be combined together to provide another cubic curve.

@@

If you you are combining $M$ color channels, each of which are values from curves of degree $N$, the resulting curve value will be degree $N+M-1$.  The intuitive explanation to this, is that each channel has the ability to add one more control point to the curve, regardless of what texture dimensionality you are working with, and each control point added raises the curve one degree.

An interesting side effect to combining color channels is that the accuracy of the resulting curve improves.  This is because you are doing some of the math in full precision floating point, instead of the limited precision of the fixed point math in the texture sampler.

@@

%-------------------------------------------------------------------------
\section*{Index of Supplemental Materials}
\label{sec:indexofsupplementalmaterials}
TODO: put information about where the zip can be downloaded.

The zip file contains a folder named WebGL which contains the WebGL demos, and a folder named GLFW which contains the GLFW (windows) C++ OpenGL demos.

\paragraph{\textbf{WebGL Folder}}
\begin{itemize}
  \item \textit{index.html} --- An index of the demos.
  \item \textit{demo.html} --- A sandbox for exploring the various options of the technique.
  \item \textit{flag.html} --- A vector art demo showing a waving flag and allowing you to modify some of the options used when rendering.
  \item \textit{rational.html} --- Rational curve demo, showing sine and cosine compared to their actual values as computed by a shader program.
\end{itemize}

\paragraph{\textbf{GLFW Folder}}
\begin{itemize}
  \item \textit{Curves} --- The folder that contains the source code and the project and solution file for the demos.
  \item \textit{glew-1.12.0} --- The glew library used for the demos.
  \item \textit{glfw-3.1.1} --- The glfw library used for the demos.  
\end{itemize}
